\documentclass{article}

\usepackage[margin=.75in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[shortlabels]{enumitem}
\usepackage{graphicx}
\usepackage{float}
\usepackage{subcaption}
\usepackage{pgfplots}

\title{Homework 2 \\ MEM 636}
\author{Athrav Joshi, Damien Prieur, Eamon Fitzgerald, Tuote Huang}
\date{}

\begin{document}

\maketitle

\section*{Problem 1}
\emph{Compute the Lie brackets for the following pair of vector fields:}
\begin{enumerate}[(1)]
\item $v_1 = \begin{bmatrix} x_2 \\ x_1 + x_2 \\ 0 \end{bmatrix}, v_2 = \begin{bmatrix} x_3 \\ x_1 \\ x_1 \end{bmatrix}$
$$ [v_1,v_2] =
\begin{bmatrix}
-x_1 \\
-x1+x_2-x_3 \\
x_2
\end{bmatrix}
$$
\item $v_1 = \begin{bmatrix} x_2 \\ x_1 + x_2 \\ 0 \end{bmatrix}, v_2 = \begin{bmatrix} x_1 \\ x_2 \\ x_3 \end{bmatrix}$
$$ [v_1,v_2] =
\begin{bmatrix}
0 \\
0 \\
0
\end{bmatrix}
$$
\end{enumerate}

\newpage
\section*{Problem 2}
\emph{Consider the nonlinear system:}
$$
\begin{cases}
\dot{x}_1   &= x_1^2x_2 \\
\dot{x}_2   &= 3x_2 + u
\end{cases}
$$
\begin{enumerate}[(1)]
\item \emph{Taylor linearize the system at the origin, $u = 0$, $x_1 = 0$, $x_2 = 0$.
            Is the linearized system controllable?}
\newline
\newline
$$
A
=
\begin{bmatrix}
\frac{\partial f_1}{\partial x_1} & \frac{\partial f_1}{\partial x_2} \\
\frac{\partial f_2}{\partial x_1} & \frac{\partial f_2}{\partial x_2}
\end{bmatrix}
\qquad
B
=
\begin{bmatrix}
\frac{\partial f_1}{\partial u} \\
\frac{\partial f_2}{\partial u}
\end{bmatrix}
$$
$$
A
=
\begin{bmatrix}
2x_1x_2 & x_1^2\\
0 & 3
\end{bmatrix}
\qquad
B
=
\begin{bmatrix}
0 \\
1
\end{bmatrix}
$$
We evaluate this at our equilibrium point, in our case the origin.
So our matrices become
$$
A
=
\begin{bmatrix}
0 & 0 \\
0 & 3
\end{bmatrix}
\qquad
B
=
\begin{bmatrix}
0 \\
1
\end{bmatrix}
\implies
\dot{x} = Ax + Bu
$$
To test for controllability we can look at the controllability matrix and look at its rank
$$
R = \begin{bmatrix} B & AB \end{bmatrix}
=
\begin{bmatrix}
0 & 0 \\
1 & 3
\end{bmatrix}
\qquad
rank(R) = 1
$$
Since the rank is not full the linearized system is not controllable.

\item \emph{Is the nonlinear system exactly feedback linearizable? If so find the normal form and linearizing control.}
\newline
\newline
For the system to be locally controllable we must show $dim \Delta_{C_0}(x_0) = n, \forall x_0 \in R^n $ where
$$ \Delta_{C_0} =  \langle f,g_1,... ,g_m | span\{g_1, ... ,g_m\}\rangle $$
$$ f = \begin{bmatrix} x_1^2x_2 \\ 3x_2\end{bmatrix} \qquad g = \begin{bmatrix}0 \\ 1 \end{bmatrix} $$
$$ [f, g](x) = \begin{bmatrix} -x_1^2 \\ -3 \end{bmatrix} \qquad [f,[f,g]] = \begin{bmatrix} 0 \\ 0 \end{bmatrix}$$
$$ \Delta_{C_0} = span\{\begin{bmatrix} x_1^2x_2 \\ 3x_2\end{bmatrix}, \begin{bmatrix} -x_1^2 \\ -3 \end{bmatrix} \} $$
Which has a dimension equal to $n$ so we are locally controllable and therefore exactlyfeedback linearizable.
Assuming the system is SISO $y = x_1$
$$z_1 = y = x_1 $$
$$z_2 = \dot{y} = \dot{x}_1 = \dot{z_1} $$
$$ \dot{z_2} = 2x_1^3x_2^2 + 3x_1^2x_2 + x_1^2u $$
$$ \text{Let, } u = \frac{-2x_1^3x_2^2-3x_1^2x_2}{x_2} + \frac{v}{x_2} $$
$u$ is our linearizing control and our linearized system is
$$ \dot{z}_1 = z_2 $$
$$ \dot{z}_2 = v $$
$$ y =  z_1 $$
Our system simplifies to
$$
\dot{\bar{z}}=
\begin{bmatrix}
0 & 1 \\
0 & 0
\end{bmatrix}
\bar{z}
+
\begin{bmatrix}
0 \\
1
\end{bmatrix}
v
$$
$$ y = \begin{bmatrix} 1 & 0 \end{bmatrix} \hat{z} $$

\item \emph{Consider the output $y = -2x_1-x_2$. Find the input-output linearizing control.
            Obtain the zero dynamics and examine their stability.
            Can the system be asymptotically stabilized?}
\newline
\newline
$$ y = -2x_1-x_2$$
$$ \dot{y} = -2\dot{x}_1 - \dot{x_2} = -2x_1^2x_2 -3x_2 - u $$
$$ \implies r = 1 \quad \alpha(x) = -2x_1^2x_2-3x_2 \quad \rho(x)= 1 \quad u = -3x_2 - 2x_1^2x_2 - v$$
$$ z_1 (x) = -2x_1 -x_2 = y $$
Choose another function to complete the transformation
$$ \xi_1(x) = x_1 $$
$$ \implies x_1 = \xi_1 \quad x_2 = -2\xi_1-z_1 $$
$$ \dot{\xi}_1(x) = x_1^2x_2  = \xi_1^2(-2\xi_1-z_1) = -2\xi_1^3 -z_1\xi_1^2$$
$$ \dot{z}_1 = v $$

$$
\begin{bmatrix}
\dot{z}_1 \\
\dot{\xi}_1
\end{bmatrix}
=
\begin{bmatrix}
0 \\
-2\xi_1^3 - z_1\xi_1^2
\end{bmatrix}
+
\begin{bmatrix}
-1 \\
0
\end{bmatrix}
v
$$
Zero Dynamics
$$ \dot{\xi}_1 = -2\xi_1^3 $$
Which is stable.
The system can be asymptotically stabilized since the internal dynamics are asymptotically stable and the output can be stabalized with a linear regulator.


\end{enumerate}

\newpage
\section*{Problem 3}
\emph{Consider a SISO system}
$$ \dot{x} = f(x) + g(x)u, \quad y=h(x) $$
\emph{Assume the system has well-defined relative degree $r$, so that the feedback linearized input-output dynamics are:}
$$ y^{(r)}=v $$
\emph{Suppose $y_R(t)$ (continuous derivatives up to $y_R^{(r)}$ will suffice) reference trajectory.
      Design a tracking controller based on stabilizing the error dynamics.
      Discuss your analysis.}
\newline
\newline
We aren't really sure how to do this one so here is an attempt.
\newline
If $y^{(r)} = v$
\newline
$y^{(r)} = L_f^rh(x) + L_gL^{r-1}h(x)u = v$
\newline
if $u = \frac{1}{L_gL^{r-1}h(x)}(-L_f^rh(x)+y_R^{(r)} + v)$
\newline
$y^{(r)} = y_R^{(r)} + v \therefore y_R^(r)$ must be $0$
\newline
$e^{(r)} = y^{(r)}-y_R^{(r)}=v $ and if $y_R^{(r)} = 0, \quad e^{(r)}=y^{(r)}=v$
\newline
This implies that the r-th derivative of the reference curve must be zero.
Any higher order derivatives will be zero and cause the algorithm to fail (in the tracking case).


\end{document}
